---
title: "Reproducible Research Assignment 1"
author: "Jon Purnell"
date: "Thursday, February 12, 2015"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
---

## Loading and preprocessing the data
```{r loadData}
options(scipen = 1)
data <- read.csv("activity.csv",header=T)
```
## What is mean total number of steps taken per day?
```{r stepsProcess}
stepsTaken <- tapply(data$steps,data$date,sum)
```

### Histogram
```{r stepsHist}
hist(stepsTaken)
```

### Mean and Median

```{r stepsStats}
stepsMean <- mean(stepsTaken,na.rm=T)
stepsMedian <- median(stepsTaken,na.rm=T)
```
Mean total steps taken is `r stepsMean`. Median total steps taken is `r stepsMedian`. 

## What is the average daily activity pattern?

A plot of the average steps taken at each 5-minute interval. 

```{r avgDaily}
avgDaily <- tapply(data$steps,data$interval,mean,na.rm=T)
plot(unique(data$interval),avgDaily,type="l",xlab="Interval",ylab="Average Daily Steps Taken")
maxInterval <- unique(data$interval)[which.max(avgDaily)]
```

The highest average of steps taken happened at `r maxInterval`. 

## Imputing missing values

```{r missing}
missingIntervals <- is.na(data$steps)
rowsMissing <- sum(missingIntervals)
```
There are `r rowsMissing` rows missing data. These are filled with the corresponding average daily steps taken for that interval.

```{r filling}
data$filledsteps <- data$steps
data$filledsteps[missingIntervals] <- avgDaily
stepsTakenFilled <- tapply(data$filledsteps,data$date,sum)
```

### Stats for filled data

```{r stepsFilledStats}
hist(stepsTakenFilled)
stepsMeanFilled <- mean(stepsTakenFilled,na.rm=T)
stepsMedianFilled <- median(stepsTakenFilled,na.rm=T)
```
Mean total steps taken is `r stepsMeanFilled`. Median total steps taken is `r stepsMedianFilled`. 

## Are there differences in activity patterns between weekdays and weekends?

```{r calculateDays, results='hide'}
weekends <- weekdays(as.Date(data$date)) %in% c("Sunday","Saturday")
data$WeekOrEnd <- factor(replace(replace(weekends,weekends == T, "Weekend"), weekends == F, "Weekday"))

par(mfrow=c(2,1))
by(data,data$WeekOrEnd,function(x) plot(unique(x$interval),tapply(x$steps,x$interval,mean,na.rm=T),type="l",xlab=x$WeekOrEnd[1],ylab="Average Daily Steps"))
```